Philosophy of Distributional Semantics\\ \large Best Practices for the Application of Computational Methods to Questions of Meaning

### 2017-05-29 pds/Paper/paper_pds.tex
get shits saved again

%\paragraph{Contribution}
%largest part of intro, preview of paper, methodology!
\paragraph{Public Attention is in ML minor extent in NLP, not in DS}
Many of the issues lately mentioned are of great importance to modern society at large. Consequently, with the rise of big data people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies. Many of these efforts have naturally been directed at the broad topic of machine learning in general.\cite{leidner2017ehtical} Only very recently the discussion has turned on machine learning used for analysis and modeling of language - often referred to as \emph{natural language processing}. Presumably because this is only one of many particular fields where machine learning is applied and the discussion of ethics natural language processing is bound pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.

Recently, however, a discussion has opened up in a hitherto completely unmentioned domain. Machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.

\paragraph{Contribution: Drawing on ML discussion for DS}
In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of distributional semantics. Specifically, we will use the \emph{Principles for Accountable Algorithms and a Social Impact Statement for Algorithms} (PAASISA) issued by the organization \emph{Fairness, Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on machine learning. 
The PAASISA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
We will assess potential violation of techniques used in distributional semantics for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.

Currently, we want to focus on aspects of discrimination and bias in the outcome of distributional semantics models for two reasons. For one, these are the least explored issues for distributional semantics. Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general machine learning debate to any sub-field that similarly employs big data and algorithms, as does computational distributional semantics. Debates concerning privacy etc. can follow the same lines as the discussion to establish general purpose ethical principles for machine learning. The issue of discrimination and bias, however, has a special relation to distributional semantics because it occupies in a unique position of computational language research. The central concern of distributional semantics, as its name gives away, is the concept of meaning. As we will argue presently, this relates it in a special way to many sensitive topics in the general machine learning discussion. Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods employed, interpretation of results and usage thereof.

@@

\paragraph{The state of the ethical discussion in ML, NLP and DS}
For the domain general use of machine learning techniques there exists a series of studies by now from legal, political computer science experts which specify principles and guidelines. Furthermore, many non-technical articles have been published outside of academic channels. 
be specific, mention FATML. use nlp ethic article to get overview.\\

>>>>

%\paragraph{Contribution}
%largest part of intro, preview of paper, methodology!
\paragraph{Public Attention is in ML minor extent in NLP, not in DS}
Many of the issues lately mentioned are of great importance to modern society at large, especially as. Consequently, people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies. Many of these efforts have naturally been directed at the broad topic of machine learning in general.\cite{leidner2017ehtical} Only a few years ago the discussion has turned on the specific use of ML for analysis and modeling of language - often referred to as \emph{natural language processing}. Presumably because this is only one of many particular fields where machine learning is applied and the discussion of ethics natural language processing is bound to pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.

A recently as 2016, however, a discussion has opened up in a hitherto completely unmentioned domain. Machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
\textbf{Why this is important. Mention Google searches using semantic information.}

The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.

\paragraph{Contribution: Drawing on ML discussion for DS}
In the present paper we want to draw on the existing efforts made by experts in the fields of ethics, computer science, public policy and law to extend these principles and guidelines to the field of distributional semantics. Specifically, we will use the \emph{Principles for Accountable Algorithms} (PAA) issued by the organization \emph{Fairness, Accountability and Transparency in Machine Learning} (FATML) - a consortium of researches and experts from private and public organizations working on machine learning. 
The PAA comprise five guiding principles: responsibility, explainability, accuracy, auditibility and fairness \cite{principles}.
We will assess potential violation of techniques used in distributional semantics for each of these principles and, where applicable, issue guidelines to prevent their negative impact upon vulnerable groups.

Currently, we want to focus on aspects of discrimination and bias in the outcome of distributional semantics models for two reasons. For one, these are the least explored ethical issues for distributional semantics. Many of the ideas, guidelines and principles concerning privacy, double use or economic impact are directly transferable from the general machine learning debate to any sub-field that similarly employs big data and algorithms. Computational distributional semantics is no exception. Debates concerning privacy etc. can follow the same lines as the discussion to establish general purpose ethical principles for machine learning. The issue of discrimination and bias, however, has a special relation to distributional semantics because the field occupies a unique position of computational language research. The central concern of distributional semantics, as its name gives away, is the concept of meaning. As we will argue presently, this relates it in a special way to many sensitive topics in the general machine learning discussion. Put succinctly, it sits at the intersection of concerns pertaining to data selection, validity or questionability of computational methods, interpretation of results and usage thereof.

@@

\paragraph{The state of the ethical discussion in ML, NLP and DS}
For the domain general use of machine learning techniques there exists a series of studies by now from legal, political computer science experts which specify principles and guidelines. Furthermore, many non-technical articles have been published outside of academic channels. 
be specific, mention FATML. \\
\textbf{use nlp ethic article to get overview. do this shits.}
